We present \texttt{micrograd}, an open-source FEniCSx framework for
density-based topology optimisation of systems governed by coupled
Brinkman--convection-diffusion equations.
The principal contributions are: (i)~a complete continuous adjoint
for a multi-objective cost functional coupling viscous dissipation
and outlet concentration mismatch; (ii)~consistent SUPG stabilisation
of both forward and adjoint transport equations at high P\'{e}clet
numbers ($\Pe\sim10^2$--$10^5$); and (iii)~a modular continuation
schedule where each component addresses a specific failure mode.

The framework is implemented in FEniCSx~(DOLFINx~v0.7.3) using
Taylor--Hood P2/P1 elements for flow and P1+SUPG elements for transport.
Robust convergence is achieved through a modular continuation schedule
--- Helmholtz filtering, Heaviside $\beta$-sharpening to $\beta=128$,
$\alpha_{\max}$ ramping, and hybrid OC/MMA ---
where each component addresses a specific diagnosed failure mode
(Table~\ref{tab:continuation}).
Functional smoothness is confirmed by a finite-difference Taylor remainder
test (slope $\approx 2.0$ on both coarse and primary meshes);
the continuously assembled adjoint provides a reliable descent direction
(Pearson correlation $\geq 0.80$ with finite-difference gradients;
OC/MMA optimisers are insensitive to gradient magnitude).

As a benchmark, the framework is applied to porous microfluidic mixing
optimisation on an $80\times20$ mesh.
The optimised permeability distribution achieves outlet concentration
RMSE of $\rmseTopOpt$ (linear target), $\rmseSigmoid$ (sigmoid), and
$\rmseGaussian$ (Gaussian peak), demonstrating framework generality
across target profiles without modifying the adjoint structure.
The gray-zone fraction is reduced to $\grayFrac$ at $\beta=\betaMax$
via $\nIterPrimary$-iteration continuation.
The optimised design exploits the continuous permeability field of the
Brinkman model to achieve precise mixing targets;
this porous optimum is not representable as a binary solid/fluid geometry
without performance penalty.
\texttt{micrograd} is deliberately a surrogate tool for rapid inverse
design exploration in porous media: it identifies high-performance
continuous-permeability concepts that inform subsequent binary design
or experimental iteration.
Threshold-projection and experimental validation are future directions.
The primary $\nIterPrimary$-iteration pipeline runs in approximately
one hour on a standard laptop; a 600-iteration run completes in under
thirty minutes with RMSE~$=\rmseShortRun$.
All results are fully reproducible via Docker and continuous integration.
