\section{Discussion}
\label{sec:discussion}

\subsection{On the choice of benchmark targets}
\label{sec:target_choice}
The linear gradient target $c^*(y)=y/L_y$ is a natural first
benchmark but is acknowledged to be relatively easy:
any geometry that partially mixes two inlet streams will
produce an approximately linear profile by diffusion.
The RMSE~$=\rmseTopOpt$ result on this target should therefore
be interpreted as a verification that the implementation
is functioning correctly, not as a demonstration of
optimisation power.

The Gaussian peak target (RMSE~$=\rmseGaussian$) is more demanding:
it requires the optimiser to route flow asymmetrically to
concentrate species near the centreline, a task that cannot
be achieved by uniform diffusion.
The sigmoid target (RMSE~$=\rmseSigmoid$) is not well-approximated
at this resolution; the result documents a limitation rather
than a capability.

More demanding benchmarks --- non-monotonic profiles, multi-peak
targets, or geometries with obstacles --- are natural extensions
but are outside the scope of this implementation paper.

\subsection{Scope and computational cost}
This paper presents a surrogate-based optimisation framework;
the contribution is the continuous adjoint derivation, SUPG-stabilised
implementation, continuation strategies, and reproducibility infrastructure,
not a directly physically realisable device design.
The Brinkman--SIMP model produces continuously graded permeability
distributions that exploit diffusive mixing through intermediate-density
porous regions; this a well-established modelling choice in
porous-media topology optimisation~\cite{borrvall2003,lazarov2016}
and is the intended physical model for this class of problems.
The design variable $\rho$ represents local permeability, not a
binary-permeability indicator; the framework optimises within the
Brinkman--SIMP design space.
The optimised design exploits the continuous permeability field
to achieve precise mixing targets; this porous optimum is not
representable as a binary solid/fluid geometry without performance penalty.
\texttt{micrograd} optimises within the Brinkman--SIMP
porous-medium model and produces continuously graded permeability
distributions that are \emph{not} directly realisable as
manufacturable geometries.
The framework is therefore a tool for studying the behaviour of
porous-medium flow-transport topology optimisation --- the
achievable RMSE within the surrogate model, the effect of
continuation strategies, and the sensitivity to mesh resolution
and volume fraction.
Users requiring binary manufacturable designs should apply
robust projection~\cite{wang2011} or iterative thresholding
during optimisation; both are identified as future work
(Section~\ref{sec:limitations}).

The primary $\nIterPrimary$-iteration optimisation on the $80\times20$ mesh
completes in approximately one hour on a standard laptop
(Intel Core~i7, 16\,GB RAM, single thread, no GPU required).
A 600-iteration run completes in under thirty minutes with
RMSE~$=\rmseShortRun$.
This is competitive with the state of the art: finite-difference
sensitivity methods require $\mathcal{O}(N)$ forward solves per
iteration (prohibitive at $N=1600$), while the continuous adjoint
requires only two adjoint solves regardless of $N$,
enabling the full $\nIterPrimary$-iteration run within one hour.

\paragraph{Computational cost breakdown.}
Each iteration requires five sparse linear system solves handled by
MUMPS direct factorisation:
\begin{itemize}\itemsep1pt
    \item Two forward solves (Brinkman + convection-diffusion):
          $\approx60\%$ of wall time.
    \item Two adjoint solves (flow adjoint + concentration adjoint):
          $\approx30\%$ of wall time.
    \item One Helmholtz filter solve: $\approx5\%$ of wall time.
    \item Assembly, sensitivity computation, OC/MMA update:
          $\approx5\%$ of wall time.
\end{itemize}
The forward and adjoint solves each factorize the same sparsity
pattern; MUMPS reuses the symbolic factorisation, reducing
the effective cost of the adjoint to approximately one
additional back-substitution per iteration.
For larger meshes ($160\times40$, 6400 elements), wall time per
iteration increases by a factor of $\approx6\times$ due to the
$\mathcal{O}(N^{1.5})$ scaling of sparse direct factorisation
in 2D; the iterative block-preconditioner (Section~\ref{sec:fem})
maintains near-optimal scaling for production meshes.

\paragraph{On framework complexity.}
The five nested continuation strategies may appear to add unnecessary
complexity. Each component, however, addresses a specific diagnosed
failure mode (Table~\ref{tab:continuation}) and the framework is
fully modular: disabling any component causes a measurable degradation
that motivates its inclusion.
Specifically: without the Helmholtz filter, checkerboard instabilities
appear by iteration~50; without $\beta$-continuation, the design
stagnates at gray fraction $>0.9$; without $\alpha_{\max}$ ramping,
early iterations converge to an all-solid minimum;
without the OC$\to$MMA switch, period-2 oscillations appear at
$\beta>8$ (Supplementary~S4).
The sinusoidal initialisation breaks the uniform-density symmetry
that otherwise prevents branching topology formation.
A user wishing to simplify the schedule for a different problem
can disable components individually by setting the corresponding
parameter to its trivial value.

\subsection{Mesh resolution and optimizer sensitivity}
\label{sec:mesh_limitation}
The optimisation is \emph{not mesh-convergent} for the
concentration RMSE objective.
The $160\times40$ mesh yields RMSE~$=\rmseHiRes$ (600 iterations)
and RMSE~$=\rmseHiResScaled$ ($\nIterHiRes$ iterations, resolution-scaled)
--- both worse than the $80\times20$ result (RMSE~$=\rmseTopOpt$).
The non-convex optimisation landscape produces different local minima
at each resolution; increasing the iteration budget does not recover
the $80\times20$ result at $160\times40$.
This is a fundamental limitation of density-based topology optimisation
applied to mixing objectives on non-uniform meshes: the continuation
schedule is not mesh-independent~\cite{lazarov2016}.
All quantitative results are therefore reported for the
$80\times20$ mesh as a \emph{proof-of-concept}.
A mesh-independent continuation schedule --- achieved by normalising
the continuation parameters by the element size $h$ --- is
the primary methodological future direction.

On coarse meshes ($20\times5$, $N_{\mathrm{out}}=6$) the outlet
discretisation is insufficient to represent the linear gradient target
and the optimizer finds unphysical reversed-flow local minima.
The $40\times10$ mesh ($N_{\mathrm{out}}=11$) converges to a
suboptimal basin; systematic improvement requires the $80\times20$
resolution or finer.
On the primary mesh, the transverse diffusion layer thickness
$\delta \approx 4\,\mu$m (Eq.~\eqref{eq:delta}) is underresolved by
a factor of $\approx 6\times$ ($\Delta y = 25\,\mu$m).
SUPG eliminates spurious oscillations but does \emph{not} recover
sub-grid diffusion structure.
The RMSE metric measures the \emph{global} outlet profile over the full
channel width $L_y = 500\,\mu$m, a scale that is well resolved;
the channel-scale branching topology is therefore correctly captured
even though the wall-normal diffusion sublayer is not.

\subsection{Positioning relative to existing frameworks}
\label{sec:positioning}
Table~\ref{tab:comparison} (Section~\ref{sec:introduction}) compares
the architectural features of \texttt{micrograd} against existing
open-source frameworks.
We highlight three specific advantages.

\paragraph{SUPG stabilisation.}
SUPG for adjoint equations is standard~\cite{donea2003,giles2000}.
We apply it to both equations using the same element-wise parameter,
which is the correct implementation~\cite{brooks1982}.
Na et al.~\cite{na2022} and Feppon et al.~\cite{feppon2024}
do not report transport stabilisation; at $\Pe\sim10^3$
this may produce spurious sensitivity-field oscillations
(Section~\ref{sec:supg_val}).
The implementation contribution here is the FEniCSx UFL
formulation for this specific coupled system at high Pe,
not the SUPG method itself.

\paragraph{FEniCSx continuous adjoint vs.\ finite differences.}
Sensitivity computation via finite differences requires $2N$
additional forward solves per iteration, where $N$ is the number
of design variables.
For the primary $80\times20$ mesh ($N=1600$), this would require
$3200$ forward solves per iteration --- computationally prohibitive.
The continuous adjoint requires only two additional adjoint solves
regardless of $N$, enabling the $\nIterPrimary$-iteration
primary run to complete in approximately one hour.
dolfin-adjoint~\cite{farrell2013} provides discrete adjoints
via automatic differentiation, which gives exact gradient magnitude
but requires the entire forward solve to be differentiable through
the FEniCSx assembly; \texttt{micrograd} trades exact magnitude
for simplicity and explicitness of the adjoint derivation.

\paragraph{Reproducibility.}
To our knowledge, \texttt{micrograd} is the first fully reproducible
open-source implementation of coupled Brinkman--convection-diffusion
topology optimisation: all results in this paper are recoverable
from a single Docker command, verified by continuous integration
on every commit.

\subsection{Limitations}
\label{sec:limitations}
\label{sec:known_limitations}
\begin{itemize}
    \item \textbf{L1: Continuous adjoint produces $L^2$ Riesz
          representative, not Euclidean gradient.}
          The assembled sensitivity vector
          $g_i = \int_\Omega (\partial J/\partial\bar\rho)\,
          \psi_i\,\mathrm{d}x$ is the $L^2$ Riesz representative
          of the gradient with respect to the $L^2$ inner product,
          not the Euclidean gradient $\hat{g} = M^{-1}g$.
          The magnitude error ($\approx 33\times$ on the primary mesh)
          does not affect OC or MMA convergence because OC depends
          on gradient sign and relative ranking~\cite{bendsoe2003},
          and MMA normalises steps by design-variable
          bounds~\cite{svanberg1987}.
          However, gradient-based line-search methods (e.g., L-BFGS)
          would require the discrete adjoint.
          A discrete adjoint via dolfin-adjoint~\cite{farrell2013}
          or automatic differentiation is identified as future work.
    \item \textbf{Porous-medium optimum and binary gap.}
          The surrogate RMSE~$=\rmseTopOpt$ is the optimum
          within the Brinkman--SIMP porous-medium model, not the
          performance of a binary solid/fluid geometry.
          Threshold-projection increases RMSE by $\binaryGapPct\%$
          (Section~\ref{sec:thresholding}), which is expected:
          the porous optimum exploits diffusive transport through
          intermediate-density regions that have no binary analogue.
          This a \emph{feature} of the surrogate approach ---
          the porous optimum reveals the thermodynamic limit of
          what is achievable with a graded-permeability medium,
          providing a design target for the binary realisation stage.
          Systematic binarisation via robust projection~\cite{wang2011}
          or iterative thresholding is deferred to future work.
    \item \textbf{Brinkman approximation.}
          All performance metrics are computed in the Brinkman--Stokes
          surrogate.
          At $\Da \approx 4\times10^{-6} \ll 1$ in solid regions, the
          surrogate accurately represents impermeability~\cite{borrvall2003}.
          Body-fitted Navier--Stokes validation is implemented in
          \texttt{ns\_validation.py} but requires \texttt{pyvista} and
          \texttt{gmsh} dependencies not present in the standard Docker
          image; these results are deferred to a companion study.
    \item \textbf{L2: The optimisation is not mesh-convergent.}
          The concentration RMSE objective does not converge
          with mesh refinement: the $160\times40$ mesh finds
          different local minima with higher RMSE at both
          600 and $\nIterHiRes$ iterations
          (Table~\ref{tab:mesh_conv}).
          All quantitative results are for the $80\times20$ mesh
          and should be treated as proof-of-concept.
          A mesh-independent continuation schedule
(normalising parameters by element size $h$) is future work.
    \item \textbf{Two-dimensional steady flow.}
          The framework is 2D; three-dimensional extension is implemented
          in \texttt{validation\_3d.py} but not systematically validated.
    \item \textbf{Fixed objective weights.}
          The weights $w_f$ and $w_c$ are fixed in this study.
          A Pareto sweep over $(w_f, w_c)$ is planned to quantify the
          trade-off between hydraulic efficiency and concentration fidelity.
    \item \textbf{No experimental validation.}
          Results are purely computational.
          Experimental fabrication and testing are necessary next steps.
\end{itemize}



\subsection{Future directions}
Immediate priorities are: (i)~a mesh-independent continuation schedule
enabling reliable optimisation at arbitrary resolution;
(ii)~adaptive mesh refinement targeting the fluid--solid interface;
(iii)~integration of body-fitted Navier--Stokes post-processing;
(iv)~a systematic Pareto front study over objective weights;
(v)~three-dimensional validation;
and (vi)~experimental collaboration for a representative subset of designs.
