\begin{thebibliography}{99}

\bibitem{jeon2000}
N.~L.~Jeon, S.~K.~W.~Dertinger, D.~T.~Chiu, I.~S.~Choi, A.~D.~Stroock, and G.~M.~Whitesides,
``Generation of solution and surface gradients using microfluidic systems,''
\emph{Langmuir}, vol.~16, no.~22, pp.~8311--8316, 2000.

\bibitem{whitesides2006}
G.~M.~Whitesides,
``The origins and the future of microfluidics,''
\emph{Nature}, vol.~442, pp.~368--373, 2006.

\bibitem{borrvall2003}
T.~Borrvall and J.~Petersson,
``Topology optimization of fluids in {S}tokes flow,''
\emph{Int. J. Numer. Methods Fluids}, vol.~41, pp.~77--107, 2003.

\bibitem{gersborg2005}
A.~Gersborg-Hansen, O.~Sigmund, and R.~B.~Haber,
``Topology optimization of channel flow problems,''
\emph{Struct. Multidiscip. Optim.}, vol.~30, pp.~181--192, 2005.

\bibitem{olesen2006}
L.~H.~Olesen, F.~Okkels, and H.~Bruus,
``A high-level programming-language implementation of topology optimization applied to steady-state {N}avier--{S}tokes flow,''
\emph{Int. J. Numer. Methods Eng.}, vol.~65, pp.~975--1001, 2006.

\bibitem{evgrafov2005}
A.~Evgrafov,
``On the limits of porous materials in the topology optimization of {S}tokes flows,''
\emph{Appl. Math. Optim.}, vol.~52, pp.~263--277, 2005.

\bibitem{challis2009}
V.~J.~Challis and J.~K.~Guest,
``Level set topology optimization of fluids in {S}tokes flow,''
\emph{Int. J. Numer. Methods Eng.}, vol.~79, pp.~1284--1308, 2009.

\bibitem{aage2008}
N.~Aage, T.~H.~Poulsen, A.~Gersborg-Hansen, and O.~Sigmund,
``Topology optimization of large scale {S}tokes flow problems,''
\emph{Struct. Multidiscip. Optim.}, vol.~35, pp.~175--180, 2008.

\bibitem{poulsen2003}
T.~H.~Poulsen,
``A new scheme for imposing a minimum length scale in topology optimization,''
\emph{Int. J. Numer. Methods Eng.}, vol.~57, pp.~741--760, 2003.

\bibitem{kreissl2010}
S.~Kreissl, G.~Pingen, A.~Evgrafov, and K.~Maute,
``Topology optimization of flexible micro-fluidic devices,''
\emph{Struct. Multidiscip. Optim.}, vol.~42, pp.~495--516, 2010.

\bibitem{kreissl2011}
S.~Kreissl, G.~Pingen, and K.~Maute,
``Topology optimization for unsteady flow,''
\emph{Int. J. Numer. Methods Eng.}, vol.~87, pp.~1229--1253, 2011.

\bibitem{kreissl2012}
S.~Kreissl and K.~Maute,
``Levelset based fluid topology optimization using the extended finite element method,''
\emph{Struct. Multidiscip. Optim.}, vol.~46, pp.~311--326, 2012.

\bibitem{alexandersen2014}
J.~Alexandersen, C.~S.~Andreasen, N.~Aage, and O.~Sigmund,
``Topology optimisation for natural convection problems,''
\emph{Int. J. Numer. Methods Fluids}, vol.~76, pp.~699--721, 2014.

\bibitem{alexandersen2020}
J.~Alexandersen and C.~S.~Andreasen,
``A review of topology optimisation for fluid-based problems,''
\emph{Fluids}, vol.~5, no.~1, p.~29, 2020.

\bibitem{makhija2015}
D.~S.~Makhija and K.~Maute,
``Level set topology optimization of scalar transport problems,''
\emph{Struct. Multidiscip. Optim.}, vol.~51, pp.~267--285, 2015.

\bibitem{deng2018}
Y.~Deng, Y.~Wu, and Z.~Liu,
\emph{Topology Optimization Theory for Laminar Flow: Applications in Inverse Design of Microfluidics}. Springer, 2018.

\bibitem{na2022}
J.~Na, H.~Li, P.~Yan, X.~Li, and X.~Gao,
``An open-source topology optimization modeling framework for the design of passive micromixer structure,''
\emph{Chem. Eng. Sci.}, vol.~260, p.~117820, 2022.

\bibitem{feppon2024}
F.~Feppon,
``Multiscale topology optimization of modulated fluid microchannels based on asymptotic homogenization,''
\emph{Comput. Methods Appl. Mech. Eng.}, vol.~419, p.~116646, 2024.

\bibitem{li2023}
H.~Li \emph{et~al.},
``A mini review on fluid topology optimization,''
\emph{Energies}, vol.~16, no.~18, p.~6497, 2023.

\bibitem{yang2020}
H.~Yang and Y.~Wang,
``Inverse design of microfluidic concentration gradient generator using deep learning and physics-based component model,''
\emph{Microfluid. Nanofluidics}, vol.~24, p.~42, 2020.

\bibitem{brooks1982}
A.~N.~Brooks and T.~J.~R.~Hughes,
``Streamline upwind/{P}etrov--{G}alerkin formulations for convection dominated flows with particular emphasis on the incompressible {N}avier--{S}tokes equations,''
\emph{Comput. Methods Appl. Mech. Eng.}, vol.~32, pp.~199--259, 1982.

\bibitem{bendsoe2003}
M.~P.~Bends{\o}e and O.~Sigmund,
\emph{Topology Optimization -- Theory, Methods, and Applications}. Springer, 2003.

\bibitem{svanberg1987}
K.~Svanberg,
``The method of moving asymptotes -- a new method for structural optimization,''
\emph{Int. J. Numer. Methods Eng.}, vol.~24, pp.~359--373, 1987.

\bibitem{elman2014}
H.~C.~Elman, D.~J.~Silvester, and A.~J.~Wathen,
\emph{Finite Elements and Fast Iterative Solvers}. Oxford University Press, 2014.

\bibitem{geuzaine2009}
C.~Geuzaine and J.-F.~Remacle,
``{G}msh: a three-dimensional finite element mesh generator with built-in pre- and post-processing facilities,''
\emph{Int. J. Numer. Methods Eng.}, vol.~79, pp.~1309--1331, 2009.

\bibitem{fenicsx2022}
M.~W.~Scroggs \emph{et~al.},
``Construction of arbitrary order finite element degree-of-freedom maps on polygonal and polyhedral cell meshes,''
\emph{ACM Trans. Math. Softw.}, vol.~48, no.~2, p.~18, 2022.


\bibitem{wang2022christmastree}
Y.-H.~Wang and Y.-S.~Sun,
``Design of Christmas-tree-like microfluidic gradient generators for cell-based studies,''
\emph{Chemosensors}, vol.~11, no.~1, p.~2, 2022.

\bibitem{kim2019finger}
J.-H.~Kim, D.-H.~Kang, B.~Kang, K.~Kim, Y.-J.~Kim, and H.~Lee,
``Finger-actuated microfluidic concentration gradient generator compatible with a microplate,''
\emph{Micromachines}, vol.~10, no.~5, p.~329, 2019.

\bibitem{fink2022automatic}
G.~Fink, T.~Mitteramskogler, M.~A.~Hinterm\"{u}ller, B.~Jakoby, and R.~Wille,
``Automatic design of microfluidic gradient generators,''
\emph{IEEE Access}, vol.~10, pp.~28155--28164, 2022.

\end{thebibliography}
