\documentclass[12pt,a4paper]{letter}
\usepackage[margin=2.5cm]{geometry}
\usepackage{hyperref}
\usepackage{microtype}

\signature{Nisong Monyimba}
\address{University of Ghana\\
Department of Mathematics\\
Accra, Ghana\\
nisongmonyimba278@gmail.com}
\date{\today}

\begin{document}

\begin{letter}{
    The Editorial Board\\
    Engineering with Computers\\
    Springer Nature
}

\opening{Dear Editors,}

We submit for your consideration the manuscript:

\begin{center}
\textbf{\texttt{micrograd}: An open-source FEniCSx framework for adjoint-based topology optimisation of porous microfluidic mixers using Brinkman--convection-diffusion equations}
\end{center}

\textbf{Summary of contributions.}
This paper presents \texttt{micrograd}, an open-source FEniCSx framework
for density-based topology optimisation of coupled
Brinkman--convection-diffusion systems.
The principal contributions are:

\begin{enumerate}
\item \textbf{Coupled continuous adjoint derivation.}
We derive the continuous adjoint for the fully coupled
Brinkman--convection-diffusion system, including the momentum--concentration
coupling term $-\lambda\nabla c$ and the Dirichlet outlet condition for the
concentration adjoint derived from the misfit functional.
This derivation is not available in the open-source FEniCSx literature for
this class of problems.

\item \textbf{SUPG-stabilised adjoint implementation.}
At the P\'{e}clet numbers relevant to microfluidic transport
($Pe \sim 10^2$--$10^5$), both the forward and adjoint transport equations
require stabilisation.
We apply SUPG consistently to both equations and verify gradient accuracy
via finite-difference Taylor test (slope $\approx 2.0$) and direction test
(Pearson correlation $\geq 0.80$ with finite-difference gradients).

\item \textbf{Continuation strategies for robust convergence.}
Five nested continuation strategies --- Helmholtz PDE filtering,
Heaviside $\beta$-continuation to $\beta=128$, logarithmic
$\alpha_{\max}$ ramping, hybrid OC/MMA scheduling, and volume-fraction
targeting --- reduce the gray-zone fraction from $0.443$ (at $\beta=16$)
to $0.074$ (at $\beta=128$).

\item \textbf{Framework generality.}
Volume-fraction sweep ($V^*\in\{0.3, 0.5, 0.7\}$) and three target
profiles (linear RMSE~$=0.058$, sigmoid RMSE~$=0.639$, Gaussian peak
RMSE~$=0.318$) demonstrate that the framework adapts to arbitrary
concentration targets without modifying the adjoint structure.

\item \textbf{End-to-end reproducibility.}
The complete pipeline is released under the MIT licence with Docker,
GitHub Actions CI, and Zenodo archival.
All results in the paper are recoverable from a single command.
\end{enumerate}

\textbf{Scope and honest limitations.}
This paper presents a \emph{surrogate-based} methodology.
The Brinkman--SIMP model produces continuously graded permeability
distributions; threshold-projection to a binary-permeability field
increases RMSE from $0.058$ to $0.359$, confirming that the optimised
design exploits gray diffusive mixing pathways that have no direct
binary physical analogue.
The paper's contribution is the numerical methodology and open-source
infrastructure.
Binary realisation, experimental validation, and a discrete adjoint
(providing exact gradient magnitude) are identified as future work.
We include a dedicated ``Known Limitations'' subsection and are committed
to transparent reporting of all negative results, including the
gradient magnitude mismatch and non-monotone mesh sensitivity.

\textbf{Suitability for IJNMF.}
This manuscript is directly suited to IJNMF's scope, which includes:
numerical methods for fluid mechanics, FEniCSx/FEniCS implementations,
PDE-constrained optimisation, and stabilised finite element methods.
The paper builds on the Borrvall--Petersson~(2003) framework published in
Engineering with Computers and extends it to coupled flow-transport
systems with a SUPG-stabilised adjoint, open-source
reproducibility infrastructure, and a modular continuation schedule ---
topics that align directly with the journal's focus on computational
engineering methodology and reproducible software frameworks.

\textbf{Suggested reviewers.}
We suggest the following potential reviewers based on their expertise
in topology optimisation, adjoint methods, and FEniCSx:
\begin{itemize}
\item Prof.\ Ole Sigmund (DTU) --- topology optimisation methodology
\item Prof.\ Boyan Lazarov (DTU/LLNL) --- Helmholtz filtering, FEniCS
\item Prof.\ Kurt Maute (CU Boulder) --- topology optimisation, adjoint methods
\item Dr.\ Florian Wechsung (NYU) --- FEniCSx adjoint methods
\end{itemize}

\textbf{Declarations.}
This manuscript has not been published elsewhere and is not under
consideration by any other journal.
All authors have approved the submission.
The code and data are publicly available at:
\url{https://github.com/nisongmonyimba278-byte/micrograd}
(Zenodo DOI: 10.5281/zenodo.XXXXXXX --- to be registered before submission).

\closing{Yours sincerely,}

\end{letter}
\end{document}
